Iterative construction of replicated designs based on Sobol' sequences

### 2016-11-23 crass/crassnew.tex
final changes (for real this time)
tex crass
Adding letter to review for CRASS and adding CRASS new lines in blue for reviewer

@@

In this note, we propose two different constructions of replicated designs of order $1$ based on Sobol' sequences. Both constructions ensure that the input space is properly explored and can be used within the replication procedure to estimate all first-order Sobol' indices. The definition of these constructions is recursive, therefore one can iteratively refine each replicated design by adding the corresponding new set of points.

We first provide a brief introduction on digital sequences and then present two iterative approaches to construct the two replicated point sets. We end this note by analyzing the space-filling properties of the two designs constructed.

@@

Consider the generating matrices $C_1,\dots,C_s$, and $C_1^m,\dots,C_s^m$ their upper left corner blocks of size $m\times m$. Although $C_1,\dots,C_s$ are of infinite size, one only requires the knowledge of $C_1^m,\dots,C_s^m$ to construct the first $2^m$ Sobol' points: for each $i=0,\dots,2^m-1$, the point $\vx_i = (x_{i,1},\dots,x_{i,s})^\intercal$ of the sequence is obtained dimension-wise by:
\begin{equation}
\label{dig.net.eq.}
(x_{i,j,1},\dots,x_{i,j,m})^\intercal = C_j^{m} \vi,\qquad j= 1,\dots,s\, ,
\end{equation}
where $x_{i,j} = \sum_{k \geq 1}^m x_{i,j,k}2^{-k}$ is the binary expansion of $x_{i,j}$ and $\vi = (i_{0},\dots,i_{m-1})^\intercal$ is the vector obtained from the binary expansion of $i= \sum_{k \geq 0}^{m-1} i_{k}2^{k}$. All matrix operations are performed in $\mathbb{F}_2$. Below we provide an example of how to compute $x_{11,1}=0.8125$ and $x_{9,2}=0.4375$,
\begin{equation*}
\begin{pmatrix}
x_{11,1,1} \\
x_{11,1,2} \\
x_{11,1,3} \\
x_{11,1,4}
\end{pmatrix} = 
\underbrace{
\begin{pmatrix}
1 & 0 & 0 & 0 \\
0 & 1 & 0 & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 0 & 1
\end{pmatrix}}_{C_1^4}
\underbrace{
\begin{pmatrix}
1 \\
1 \\
0 \\
1
\end{pmatrix}}_{\bvec{11}}=
\begin{pmatrix}
1 \\
1 \\
0 \\
1
\end{pmatrix}
,\qquad
\begin{pmatrix}
x_{9,2,1} \\
x_{9,2,2} \\
x_{9,2,3} \\
x_{9,2,4}
\end{pmatrix} = 
\underbrace{\begin{pmatrix}
1 & 1 & 1 & 1 \\
0 & 1 & 0 & 1 \\
0 & 0 & 1 & 1 \\
0 & 0 & 0 & 1
\end{pmatrix}}_{C_2^4}
\underbrace{\begin{pmatrix}
1 \\
0 \\
0 \\
1
\end{pmatrix}}_{\bvec{9}} =
\begin{pmatrix}
0 \\
1 \\
1 \\
1
\end{pmatrix}
\end{equation*}

@@

\begin{lemma}
Under the additive construction $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ are two replicated designs of order $1$, $\ell \geq 0$.
\end{lemma}
\textit{Proof}: For each $\ell\geq 1$, digital shift and scrambling operations carried out in equations (\ref{linear.scrambl}) and (\ref{dig.shift}) are bijections from $\mathbb{F}_2^r$ and $\left(\mathbb{F}_2^{r}\right)^{s}$. Hence, the same coordinates of $B_\ell$ and ${B'}_\ell$ will contain the same set of $2^r$ values.

@@

We name by design $M$ (resp. $A$) one of the two replicated designs constructed with the multiplicative (resp. additive) approach, and set the input space dimension $s$ to $6$. For design $M$, we study the properties of $\mathcal{P}_8,\dots,\mathcal{P}_{12}$, and for design $A$, we set the initial value to $r=8$ and study the properties of $\mathcal{P}_0,\dots,\mathcal{P}_{15}$. For both $M$ and $A$ we generate $100$ designs whose average criteria are presented in Figure \ref{results.crit}. While each of the $100$ designs $M$ is independently randomized applying Owen's scrambling \cite{OwenScrambling}, designs $A$ are constructed by randomly selecting matrix $L$ (uniformly among all $2^{r(r-1)/2}$ choices) and vectors $\ve_\ell$ used in equation (\ref{dig.shift}). In addition, we also show in Figure \ref{results.crit} the results obtained with a Latin Hypercube design (LH design) with an equal number of points as in design $A$. LH designs are the standard designs used in the replication procedure to estimate first-order indices. The x-axis corresponds to the number of design points $N$ and the y-axis to the value of the criterion.

As expected, design $M$ performs the best and the estimated slope of the discrepancy curve ($-0.72$) falls within the expected range for Sobol' sequences \cite{Morokoff}. Although design $A$ does not perform as well as design $M$, it outperforms a classic LH design for both criteria while allowing $15$ refinement steps from $2^8$ to $2^{12}$ points, instead of only $4$ steps for design $M$.
\begin{figure}
\caption{Log-log graph of averaged $L_2$ star discrepancy (left) and maximin (right) for designs $M$ and $A$ over $100$ repetitions.}
\label{results.crit}
\centering
\includegraphics[scale=0.38]{discrep.pdf}
\hspace*{0.1cm}
\includegraphics[scale=0.38]{mindist.pdf}
\end{figure}

@@

% The Acknowledgements are an un-numbered section
%\section*{Acknowledgements}
% Acknowledgements text here

@@

\bibitem{Matousek} J. Matou$\check{s}$ek, On the $L2$-discrepancy for anchored boxes, \emph{Journal of Complexity} 14, 1998, pp. 527?-556.

>>>>

\usepackage{color} % TO REMOVE AFTER REVIEW APPROVES COMMENTS IN BLUE

@@

In this note, we propose two different constructions of replicated designs of order $1$ based on Sobol' sequences. Both constructions ensure that the input space is properly explored and can be used within the replication procedure to estimate all first-order Sobol' indices. The definition of these constructions is recursive, therefore one can iteratively refine each replicated design by adding the corresponding new set of points. We first provide a brief introduction on digital sequences and then present two iterative approaches to construct the two replicated point sets. We end this note by analyzing the space-filling properties of the two designs constructed.

@@

Consider the generating matrices $C_1,\dots,C_s$, and $C_1^m,\dots,C_s^m$ their upper left corner blocks of size $m\times m$. Although $C_1,\dots,C_s$ are of infinite size, one only requires the knowledge of $C_1^m,\dots,C_s^m$ to construct the first $2^m$ Sobol' points: for each $i=0,\dots,2^m-1$, the point $\vx_i = (x_{i,1},\dots,x_{i,s})^\intercal$ of the sequence is obtained dimension-wise by:
\begin{equation}
\label{dig.net.eq.}
(x_{i,j,1},\dots,x_{i,j,m})^\intercal = C_j^{m} \vi,\qquad j= 1,\dots,s\, ,
\end{equation}
where $x_{i,j} = \sum_{k \geq 1}^m x_{i,j,k}2^{-k}$ is the binary expansion of $x_{i,j}$ and $\vi = (i_{0},\dots,i_{m-1})^\intercal$ is the vector obtained from the binary expansion of $i= \sum_{k \geq 0}^{m-1} i_{k}2^{k}$. All matrix operations are performed in $\mathbb{F}_2$. Below we provide an example of how to compute $x_{11,1}=0.8125$ and $x_{9,2}=0.4375$,
\begin{equation*}
\renewcommand{\arraystretch}{1.3}
\begin{pmatrix}
x_{11,1,1} \\
x_{11,1,2} \\
x_{11,1,3} \\
x_{11,1,4}
\end{pmatrix} = 
\underbrace{
\begin{pmatrix}
1 & 0 & 0 & 0 \\
0 & 1 & 0 & 0 \\
0 & 0 & 1 & 0 \\
0 & 0 & 0 & 1
\end{pmatrix}}_{C_1^4}
\underbrace{
\begin{pmatrix}
1 \\
1 \\
0 \\
1
\end{pmatrix}}_{\bvec{11}}=
\begin{pmatrix}
1 \\
1 \\
0 \\
1
\end{pmatrix}
,\qquad
\begin{pmatrix}
x_{9,2,1} \\
x_{9,2,2} \\
x_{9,2,3} \\
x_{9,2,4}
\end{pmatrix} = 
\underbrace{\begin{pmatrix}
1 & 1 & 1 & 1 \\
0 & 1 & 0 & 1 \\
0 & 0 & 1 & 1 \\
0 & 0 & 0 & 1
\end{pmatrix}}_{C_2^4}
\underbrace{\begin{pmatrix}
1 \\
0 \\
0 \\
1
\end{pmatrix}}_{\bvec{9}} =
\begin{pmatrix}
0 \\
1 \\
1 \\
1
\end{pmatrix}
\end{equation*}

@@

\begin{lemma}
Under the additive construction $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ are two replicated designs of order $1$, $\ell \geq 0$.
\end{lemma}
\textit{Proof}: For each $\ell\geq 1$, digital shift and scrambling operations carried out in (\ref{linear.scrambl}) and (\ref{dig.shift}) are bijections from $\mathbb{F}_2^r$ and $\left(\mathbb{F}_2^{r}\right)^{s}$. Hence, the same coordinates of $B_\ell$ and ${B'}_\ell$ will contain the same set of $2^r$ values.

@@

We name by design $M$ (resp. $A$) one of the two replicated designs constructed with the multiplicative (resp. additive) approach, and set the input space dimension $s$ to $6$. For design $M$, we study the properties of $\mathcal{P}_8,\dots,\mathcal{P}_{12}$, and for design $A$, we set the initial value to $r=8$ and study the properties of $\mathcal{P}_0,\dots,\mathcal{P}_{15}$. For both $M$ and $A$ we generate $100$ designs whose average criteria are presented in Figure \ref{results.crit}. While each of the $100$ designs $M$ is independently randomized applying Owen's scrambling \cite{OwenScrambling}, designs $A$ are constructed by randomly selecting matrix $L$ (uniformly among all $2^{r(r-1)/2}$ choices) and vectors $\ve_\ell$ used in equation (\ref{dig.shift}). \textcolor{blue}{In addition, we also show in Figure \ref{results.crit} the results obtained with an optimized Latin Hypercube design (LH design) according to each criterion for an equal number of points as in design $A$.} The x-axis corresponds to the number of design points $N$ and the y-axis to the value of the criterion. \textcolor{blue}{Although LH designs are the standard designs used in the replication procedure to estimate first-order indices, the replicated LH design of an optimized LH design is not guaranteed to be optimal.}

As expected, design $M$ performs the best and the estimated slope of the discrepancy curve ($-0.72$) falls within the expected range for Sobol' sequences \cite{Morokoff}. \textcolor{blue}{Although design $A$ does not perform as well as design $M$, it outperforms an optimized LH design for the discrepancy criterion. However for the maximin criterion, design $A$ shows slightly worse results. Concerning the number of steps, design $A$ allows $15$ refinement steps from $2^8$ to $2^{12}$ points, instead of only $4$ steps for design $M$.}
\begin{figure}
\caption{Log-log graph of averaged $L_2$ star discrepancy (left) and maximin (right) for designs $M$ and $A$ over $100$ repetitions.}
\label{results.crit}
\centering
\includegraphics[scale=0.38]{discrep.pdf}
\hspace*{0.1cm}
\includegraphics[scale=0.38]{mindist.pdf}
\end{figure}

@@

% The Acknowledgements are an un-numbered section
\section*{Acknowledgements}
This work is supported by the CITiES project funded by the Agence Nationale de la Recherche (grant ANR-12-MONU-0020).

@@

\bibitem{Matousek} J. Matou$\check{s}$ek, On the $L2$-discrepancy for anchored boxes, \emph{Journal of Complexity} 14, 1998, pp. 527-556.
